---
title: "Geog 210B - Lab 2"
author: "Conor McMahon"
date: "1/27/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)    # data frame handling
library(janitor)      # cleaning data frame format
library(kableExtra)   # table styling
library(gtools)       # combinatorix

```


### 1) The probability of a dry summer is equal to 0.3, the probability of a wet summer is equal to 0.2, and the probability of a summer with normal precipitation is equal to 0.5. A climatologist observed the precipitation during three consecutive summers.

#### a) Enumerate the sample space, and assign probabilities to each simple event. Do they add to one?

With three summers that can be one of three outcomes (wet, dry, or normal), there are *3^3^ = 27* potential outcomes. If we call the wet/dry/normal character the random variable *prec*, and choose *prec* three times, the possible perumations are:

```{r echo=F}

prec_outcomes <- c('dry','wet','normal')
prec_perm <- permutations(n=3, r=3,  # "3 choose 3"
                                  v=prec_outcomes, repeats.allowed=T)

prec_perm_df <- data.frame(First_Summer  = prec_perm[,1],
                           Second_Summer = prec_perm[,2],
                           Third_Summer  = prec_perm[,3],
                           Num_Wet       = c(0,0,1,0,0,1,1,1,2,0,0,1,0,0,1,1,1,2,1,1,2,1,1,2,2,2,3),
                           Num_Dry       = c(3,2,2,2,1,1,2,1,1,2,1,1,1,0,0,1,0,0,2,1,1,1,0,0,1,0,0))
# Build a string-coded probability density function
prob_map <- new.env(hash = TRUE, parent = emptyenv(), size = NA)
prob_map[['dry']] <- 0.3
prob_map[['wet']] <- 0.2
prob_map[['normal']] <- 0.5

prec_perm_df$Probability = prob_map[prec_perm_df$First_Summer]*
                           prob_map[prec_perm_df$Second_Summer]*
                           prob_map[prec_perm_df$Third_Summer]

kable(prec_perm_df) %>% kable_styling()

```

Counting the outcomes above, we arrive at the following enumeration of the sample space (list of combinations and expected frequencies):

```{r}

prec_comb <- combinations(n=3, r=3, # "3 choose 3"
                          v=prec_outcomes, repeats.allowed=T)

prec_perm_df <- data.frame(Num_Wet   = c(0,0,1,0,1,2,0,1,2,3),
                           Num_Dry   = c(3,2,2,1,1,1,0,0,0,0)
                           )


```




### b. What is the probability of observing two dry summers?

### c. What is the probability of observing at least two dry summers?

### d. What is the probability of not observing a wet summer?